\hypertarget{structure-organization}{%
 \section{STRUCTURE \& ORGANIZATION}
 \label{structure-organization}}

\hypertarget{the-executive}{%
 \subsection{The Executive}
 \label{the-executive}}

The MES Executive shall consist of the following positions. One person
shall occupy each position.

\import{./}{president.tex}
\import{./}{vice-president-student-life.tex}
\import{./}{vice-president-external-relations.tex}
\import{./}{vice-president-academic.tex}
\import{./}{vice-president-finance.tex}
\import{./}{vice-president-internal.tex}
\import{./}{vice-president-communications.tex}

\hypertarget{authority-and-responsibility-of-the-executive}{%
 \subsection{Authority and Responsibility of the
  Executive}
 \label{authority-and-responsibility-of-the-executive}}
\hypertarget{executive-authority}{%
 \subsubsection{Executive Authority}
 \label{executive-authority}}

The Executive has the following authority:

\begin{enumerate}
 \item
  To administer the day-to-day business and operations of the MES.
 \item
  To act as an advisory board to the MES Council on the effects of the
  MES Council's policy decisions on the day-to-day operations of the
  MES.
 \item
  To perform the functions of the MES Council during the summer, subject
  to any restrictions placed on it by the MES Council.
 \item
  In case of an emergency situation where the MES Council is unable to
  meet, the Executive may make an interim decision, although any policy
  decisions shall be subject to ratification at the next MES Council
  meeting.
 \item
  Approve expenditures not exceeding \$500, unless the expenditure is
  coming from the Executive Operations account.

\end{enumerate}

\hypertarget{executive-responsibility}{%
 \subsubsection{Executive
  Responsibility}
 \label{executive-responsibility}}
The Executive shall have the following responsibility:

\begin{enumerate}
 \item
  To review the MES budget prepared by the VPF for final recommendations
  to the MES Council.
 \item
  To form ad-hoc committees when required for the purpose of appointing
  members to assist in administering the policies of the MES and
  undertaking new initiatives.
 \item
  To be accountable and transparent for all its actions to the MES
  Council and to the general membership at each General Meeting.
 \item
  Initiate and encourage new events and initiatives of the MES.
 \item
  Supervise appointed MES Council positions to provide advice or
  assistance.
 \item
  Liaise with other members or groups within the university community.
 \item
  Promote professionalism within the MES Council.
 \item
  Take a proactive role in the long term goal setting of the MES.
 \item
  Ensure effective transition and continuity between the outgoing
  Council and the incoming Council, through documentation and
  orientation.
 \item
  Create a positive and effective working environment.
 \item
  Executive members to hold at least 2 hours of scheduled office hours
  weekly.

\end{enumerate}

\hypertarget{associate-vice-presidents}{%
 \subsection{Associate
  Vice-Presidents}
 \label{associate-vice-presidents}}

\hypertarget{associate-vice-president-events}{%
 \subsubsection{Associate Vice President,
  Events}
 \label{associate-vice-president-events}}

The Associate Vice President, Events shall:

\begin{enumerate}
 \item
  Oversee the organization of social events of the MES foremost, but not
  limited to:

  \begin{enumerate}
   \item
    Eng Pub Nights
   \item
    Fireball
   \item
    Kipling
   \item
    Sports
  \end{enumerate}
 \item
  Chair the Social Committee (see MES Bylaws Section D.4.11).
 \item
  Initiate at least one new activity and ensure the overall schedule of
  events offers a variety of activities that will appeal to many
  different groups.
 \item
  Coordinate with the VPC to promote MES activities.
 \item
  Provide the Administrator with dates and details of all events and
  initiatives.
 \item
  Report to Vice President, Student Life.
 \item
  Supervise the following positions:

  \begin{enumerate}
   \item
    Fireball Coordinator(s)
   \item
    Kipling Coordinator(s)

  \end{enumerate}
\end{enumerate}

\hypertarget{associate-vice-president-clubs}{%
 \subsubsection{Associate Vice President,
  Clubs}
 \label{associate-vice-president-clubs}}
The Associate Vice President, Clubs shall:

\begin{enumerate}
 \item
  Supervise the care of all MES rental equipment.
 \item
  Prepare an interim progress report and transition report (see Appendix
  J).
 \item
  Recommend funding levels for all MES funded Clubs and Teams to the
  Vice President, Finance and Administration and enforce the Funding
  Policy (see MES Bylaws Section H.3).
 \item
  Assist Vice President, Internal in the allocation of Hatch Space to
  Clubs and Teams demonstrating essential need.
 \item
  Promote MES flagship by encouraging the use of MES branding amongst
  MES groups.
 \item
  Chair the Club Leaders Committee (see MES Bylaws Section D.4.3).
 \item
  Assist Vice President, Student Life with the organization of
  Back2School Week.
 \item
  Report to Vice President, Internal.

\end{enumerate}

\hypertarget{associate-vice-president-academic-resources}{%
 \subsubsection{Associate Vice President, Academic
  Resources}
 \label{associate-vice-president-academic-resources}}
The Associate Vice President, Academic Resources shall:

\begin{enumerate}
 \item
  Organize all academically related events run by the MES including, but
  not limited to:

  \begin{enumerate}
   \item
    MES test preparation sessions
   \item
    Technical skills sessions
  \end{enumerate}
 \item
  Be responsible for maintaining all aspects of the McMaster Engineering
  Society Textbook Library, foremost:

  \begin{enumerate}
   \item
    Actively working to procure new textbooks
   \item
    Maintaining an updated inventory of all textbooks in the MES'
    possession.
   \item
    Coordinating with the Drain Coordinators to ensure proper
    instruction is given to the Drain volunteers on how to run the
    Textbook Library.
   \item
    Cleaning the Textbook Library and ensuring it is kept organized.
   \item
    Ensuring all textbooks in the Textbook Library are labelled with the
    MES logo.
   \item
    Ensuring all textbooks are returned to the Textbook Library by the
    end of each semester.
  \end{enumerate}
 \item
  Work with the VPA to plan Back2School Week
 \item
  Chair the Academic Services Committee (See MES Bylaws Section D.4.18)
 \item
  Actively seek ways to improve the academically related events and
  services provided by the MES.
 \item
  Organize and maintain the Notes/Test Bank.
 \item
  Report to Vice President, Academic.
\end{enumerate}

\hypertarget{representatives}{%
 \subsection{Representatives}
 \label{representatives}}
\hypertarget{program-representatives}{%
 \subsubsection{Program
  Representatives}
 \label{program-representatives}}

\begin{enumerate}
 \item
  There shall be one representative from each program listed below, who
  is a member of that program.

  \begin{enumerate}
   \item
    Chemical Engineering
   \item
    Civil Engineering
   \item
    Software Engineering
   \item
    Electrical and Computer Engineering
   \item
    Engineering Physics
   \item
    Materials Science and Engineering
   \item
    Mechanical Engineering
   \item
    Mechatronics Engineering
   \item
    Computer Science
   \item
    Engineering and Management
   \item
    Engineering and Society
   \item
    Bachelor of Technology
   \item
    Integrated Biomedical Engineering and Health Sciences
  \end{enumerate}

 \item
  Responsibilities:

  \begin{enumerate}
   \item
    To act as a liaison between the MES, their respective program
    societies, and the students represented by those societies.
   \item
    To keep the MES informed of all program and program society issues
    and activities.
   \item
    Advertise all MES events and activities to their respective
    constituents by making class announcements, distributing posters
    from the Graphic Designers, distributing MES publications, promoting
    the MES website, and MESsenger email list.
   \item
    Actively encourage participation in all MES activities.
   \item
    Ensure quorum is met at General Meetings
   \item
    Serve as a representative for their program on the macLAB Board of
    Directors.
   \item
    Organize their program's team(s) for MES Sports Tournaments, with
    the exception of Management and Society.
   \item
    Represent the MES at their respective Program Society's meetings.
   \item
    Report to the President.
  \end{enumerate}

\end{enumerate}

\hypertarget{first-year-representatives}{%
 \subsubsection{First Year
  Representatives}
 \label{first-year-representatives}}
\begin{enumerate}
 \item
  Six Level I Representatives, who are full members of the MES, will be
  elected from the first year class that they represent with the
  following number of representatives from each program:

  \begin{enumerate}
   \item
    Three from Engineering 1
   \item
    One from Computer Science
   \item
    One from Bachelor of Technology
   \item
    One from iBioMed
  \end{enumerate}
 \item
  Responsibilities:

  \begin{enumerate}
   \item
    Advertise all MES events and activities to first year students by
    making class announcements, promoting the MES website, and MESsenger
    email list.
   \item
    Serve on the First Year Committee (see ``Committees'', Section
    I.4.8).
   \item
    Represent the concerns and opinions of first year students within
    the Faculty of Engineering.
   \item
    Promote and encourage first year attendance at all events.
   \item
    Promote MES committees and activities to the first year students.
   \item
    Seek out course coordinators to voice concerns of the first year
    students.
   \item
    Assist the Engineering Student Services office with their first year
    student events.
   \item
    Assist in the recruitment and transition of subsequent First Year
    Representatives.
   \item
    Organize first year student team(s) for the Ratboy Soccer
    Tournament.
   \item
    Report to the VPSL.
   \item
    If applicable, represent the MES at their respective program
    society's meetings.
  \end{enumerate}
\end{enumerate}

\hypertarget{executive-transition-process}{%
 \subsection{Executive Transition
  Process}
 \label{executive-transition-process}}
\begin{enumerate}
 \item
  The outgoing and incoming MES Executive shall plan and attend a
  retreat to make long term plans for the year ahead and carry out a
  more detailed transition.
 \item
  The budget for this retreat shall be set for no more than \$2000,
  taken from the Executive Operations budget line.

  \begin{enumerate}
   \item
    The VPSL is to organize, plan and document the retreat.
  \end{enumerate}
 \item
  The incoming MES Executive shall prepare a presentation for the first
  MES Council meeting of their term, outlining what was accomplished on
  the retreat, what the budget was used for, and informing the MES
  Council of the year's objectives and long-term goals.
\end{enumerate}

\hypertarget{elections-selections}{%
 \section{\texorpdfstring{\emph{ELECTIONS \&
     SELECTIONS}}{ELECTIONS \& SELECTIONS}}
 \label{elections-selections}}
\hypertarget{general}{%
 \subsection{General}
 \label{general}}
\begin{enumerate}
 \item
  There shall be two classes of internal MES elections: regular
  elections and by-elections.
 \item
  All MES elections shall be administered and overseen by the Elections
  Committee (see MES Bylaws Section D.4.6).
 \item
  Election dates will be determined at the discretion of the Elections
  Committee.
 \item
  At least 3\% of all eligible McMaster undergraduate students in the
  Faculty of Engineering must vote for an election to be considered
  valid.
 \item
  Elections shall be held online and officiated by the Faculty of
  Engineering.

\end{enumerate}

\hypertarget{regular-elections}{%
 \subsection{Regular Elections}
 \label{regular-elections}}
\begin{enumerate}
 \item
  A meeting mandatory for all candidates will be held by the Chief
  Returning Officer prior to the campaign period of each regular
  election to inform candidates of the campaign rules and important
  dates as determined by the Elections Committee.

  \begin{enumerate}
   \item
    The consequences for an unexcused absence from this meeting on the
    part of a candidate shall be decided on a case-by-case basis by the
    Elections Committee but may include removal from the election if the
    Elections Committee sees fit.
  \end{enumerate}
 \item
  There shall be four distinct types of regular elections: presidential
  elections, general elections, Bachelor of Technology elections, and
  First Year Representative elections.
 \item
  A ``meet-and-greet'' style event will be held by the Chief Returning
  Officer during the campaign period of every regular election, except
  for First Year Representative elections.

\end{enumerate}

\hypertarget{presidential-elections}{%
 \subsubsection{Presidential
  Elections}
 \label{presidential-elections}}
\begin{enumerate}
 \item
  The presidential elections shall serve to elect the MES President.
 \item
  Presidential elections are to be held prior to Fireball, and the
  winning presidential candidate shall be announced at Fireball.
 \item
  All McMaster undergraduate students in the Faculty of Engineering
  shall be eligible to vote for the MES President.

\end{enumerate}

\hypertarget{bachelor-of-technology-election}{%
 \subsubsection{Bachelor of Technology
  Election}
 \label{bachelor-of-technology-election}}
\begin{enumerate}
 \item
  The Bachelor of Technology elections shall serve to elect the Bachelor
  of Technology Representative.
 \item
  Bachelor of Technology elections are to be held in November, with
  polls closing prior to the last day of Fall term classes.
 \item
  Only the students currently enrolled in the Bachelor of Technology
  program shall be eligible to vote for the Bachelor of Technology
  Representative.
\end{enumerate}

\hypertarget{general-elections}{%
 \subsubsection{General Elections}
 \label{general-elections}}
\begin{enumerate}

 \item
  The following MES Council members shall be elected at the time of the
  a general election:

  \begin{enumerate}
   \item
    Vice President, Student Life
   \item
    Vice President, External Relations
   \item
    Vice President, Academic
   \item
    Vice President, Finance
   \item
    Vice President, Communications
   \item
    Vice President, Internal
   \item
    Associate Vice President, Events
   \item
    Associate Vice President, Clubs
   \item
    Associate Vice President, Academic Resources
   \item
    Chemical Engineering Representative
   \item
    Civil Engineering Representative
   \item
    Computer Science Representative
   \item
    Electrical and Computer Engineering Representative
   \item
    Engineering and Management Representative
   \item
    Engineering and Society Representative
   \item
    Engineering Physics Representative
   \item
    Integrated Biomedical Engineering and Health Sciences Representative
   \item
    Materials Science and Engineering Representative
   \item
    Mechanical Engineering Representative
   \item
    Mechatronics Engineering Representative
   \item
    Software Engineering Representative
  \end{enumerate}
 \item
  The general election shall be completed a week before the second term
  General Meeting at the latest.
 \item
  All McMaster undergraduate students in the Faculty of Engineering
  shall be eligible to vote for all the positions listed above except
  for the Program Representatives. Only the aforementioned students
  currently enrolled in the corresponding program shall be eligible to
  vote for their Program Representatives.

\end{enumerate}

\hypertarget{first-year-representative-elections}{%
 \subsubsection{First Year Representative
  Elections}
 \label{first-year-representative-elections}}
\begin{enumerate}
 \item
  Six First Year Representatives shall be elected from the first year
  class of McMaster undergraduate students in the Faculty of
  Engineering.

  \begin{enumerate}
   \item
    Only first year students enrolled in the McMaster Faculty of
    Engineering are eligible to vote. These students shall only be
    eligible to vote for the representatives of the program that they
    are enrolled in
  \end{enumerate}
 \item
  There shall be a nomination period of one week during which all
  interested first year students may submit their nomination forms to
  the Chief Returning Officer.
 \item
  Elections shall be held in September.
\end{enumerate}
\hypertarget{by-elections}{%
 \subsection{By-Elections}
 \label{by-elections}}

\begin{enumerate}
 \item
  By-elections may be held to fill a vacancy in an elected office due to
  a reason other than the expiration of the term of office of the
  official in question.

  \begin{enumerate}
   \item
    A by-election need not be held to fill a vacancy if a position
    becomes vacant within the last eight weeks of the term of office of
    the elected official.
  \end{enumerate}
 \item
  By-elections will be conducted with their own set of rules as follows:

  \begin{enumerate}
   \item
    The nomination period shall last a minimum of 5 days (including
    weekend days)
   \item
    The campaign period shall last a minimum of 5 weekdays
   \item
    Polling shall immediately follow the campaign period
  \end{enumerate}
\end{enumerate}
\hypertarget{eligibility-for-elected-positions}{%
 \subsection{Eligibility for Elected
  Positions}
 \label{eligibility-for-elected-positions}}

\begin{enumerate}
 \item
  Candidates are not required to be MES members at the time of their
  campaign, but if elected they must become full members of the MES as
  detailed in the Constitution and must remain members for the duration
  of their term.

  \begin{enumerate}
   \item
    Individuals may hold an elected position on the MES without holding
    an MES membership throughout their term. This exception shall be
    granted on a financial need basis at the discretion of the Chief
    Returning Officer.
  \end{enumerate}
 \item
  Candidates for President must be registered in Level III or higher at
  the time of election.
 \item
  Candidates for VPSL, VPA, VPX, VPC, VPI, and VPF must be enrolled in
  Level II or higher at the time of election.
 \item
  Candidates for Program Representatives must be enrolled in the program
  they are campaigning for at the time of election

  \begin{enumerate}
   \item
    In the case a program representative changes their enrolment program
    or resigns from their position, the seat will be immediately
    vacated. The Elections Committee will determine the election process
    for the vacated position.
  \end{enumerate}
 \item
  In the case of a by-election for a Vice-President or Program
  Representative during the academic year, candidates must be enrolled
  in Level II or higher at the time of election.
 \item
  Candidates for First Year Representative must be registered in Level I
  at the time of the election or by-election for that position.
 \item
  All candidates must follow the rules set down in these bylaws, or in
  any other duly enacted document, and any rulings issued by the
  Elections Committee.
 \item
  Eligibility exceptions can be made at the discretion of the Elections
  Committee.
\end{enumerate}

\hypertarget{eligibility-for-appointed-positions}{%
 \subsection{Eligibility for Appointed
  Positions}
 \label{eligibility-for-appointed-positions}}

\begin{enumerate}
 \item
  Chief Returning Officer candidates must have held at least one voting
  or appointed position on the MES Council for a complete term.
 \item
  At least one Culture Coordinator(s) must be registered in Level III or
  higher at the time of their appointment.
 \item
  Co-Orientation Coordinators must meet eligibility requirements, as
  outlined in C.3.3.a)

\end{enumerate}

\hypertarget{election-dates}{%
 \subsection{Election Dates}
 \label{election-dates}}

\hypertarget{election-dates-1}{%
 \subsubsection{Election Dates}
 \label{election-dates-1}}
All Society elections shall consist of three distinct periods of time:
the nomination period, the campaign period and the election period.

\begin{enumerate}
 \item
  The nomination period for all MES elections shall extend over at least
  10 full days, during which classes are in session. Duly completed
  nomination forms shall only be accepted by the Chief Returning Officer
  during this period.
 \item
  The campaign period for the presidential and general elections shall
  span a period of no less than 5 full weekdays during which classes are
  held. The weekend days enclosed by the campaign commencement and
  termination dates will be considered open for campaigning.
 \item
  It is recommended that at least one day be scheduled between the close
  of nomination period and the commencement of campaign period.
 \item
  The First Year Representative elections' campaign period shall span a
  minimum of 4 full weekdays during which classes are held. The weekend
  days enclosed by the campaign commencement and termination dates will
  be considered open for campaigning.
 \item
  All casting of ballots for all MES elections shall be conducted
  immediately following the campaign period on a day in which classes
  are held. The McMaster undergraduate students in the Faculty of
  Engineering of the appropriate constituencies shall have an
  opportunity to cast ballots over a period of no less than five
  consecutive hours on the election day.

  \begin{enumerate}
   \item
    For online polling, the balloting period must extend for a minimum
    of eighteen consecutive hours.
  \end{enumerate}
 \item
  Election dates are set at the discretion of the Elections Committee.

\end{enumerate}
\hypertarget{term-of-office}{%
 \subsection{Term of Office}
 \label{term-of-office}}
\begin{enumerate}
 \item
  The term of office for all MES Council members, with the exception of
  the First Year and Bachelor of Technology Representatives, shall
  commence the day of the General Meeting, to be held in March of the
  year of election, upon ratification and terminate the day of the
  General Meeting to be held in March of the following year, upon
  ratification of the incoming MES Council.
 \item
  The term of office of the First Year Representatives shall commence at
  the first MES Council meeting following their election, pending
  ratification by the MES Council, and terminate on the last day of
  class of that academic year.
 \item
  The term of office of the Bachelor of Technology Representative shall
  commence at the first MES Council meeting after the conclusion of
  their election, pending ratification by the MES Council, and shall
  terminate after the first council meeting back after the winter break.
 \item
  An MES Council member cannot hold more than one voting position, on
  the MES Council, simultaneously. Should this occur, the MES Council
  member in question must resign from one of the positions.

\end{enumerate}
\hypertarget{nominations}{%
 \subsection{Nominations}
 \label{nominations}}
\begin{enumerate}
 \item
  All eligible McMaster undergraduate students in the Faculty of
  Engineering (see MES Bylaws Section C.1.1) wishing to run for any MES
  Council position open for election shall present an appropriate
  nomination form (see Appendix G) signed by the relevant amount of
  McMaster undergraduate students in the Faculty of Engineering, of the
  constituency in question, to the Chief Returning Officer during the
  nomination period.
  \begin{enumerate}
   \item
    All executive and associate vice president positions (President,
    VPs, AVPs) are required to have a minimum of 20 nomination
    signatures.
   \item
    All elected program representative positions are required to have a
    minimum of 10 nomination signatures.
  \end{enumerate}
 \item
  If, at the conclusion of the nomination period there are no
  nominations for an MES Council position open for election, the Chief
  Returning Officer will make arrangements for by-elections to fill the
  vacancy.

  \begin{enumerate}
   \item
    If, at the conclusion of the nomination period there is only one
    nomination for any MES Council position, an election for that
    position must still be held.
  \end{enumerate}
 \item
  It is the responsibility of the Chief Returning Officer to ensure the
  publication of elected position nominees, in a timely fashion. This
  may be published to a page on the MES website, the MES Facebook page,
  or to another public forum accessible to any McMaster undergraduate
  students in the Faculty of Engineering.

\end{enumerate}

\hypertarget{campaign}{%
 \subsection{Campaign}
 \label{campaign}}
\hypertarget{campaign-conduct}{%
 \subsubsection{Campaign Conduct}
 \label{campaign-conduct}}
\begin{enumerate}
 \item
  All candidates must conduct their campaign according to the campaign
  rules as determined and set out by the Elections Committee.
 \item
  Any violation of these rules will be reviewed by the Elections
  Committee and subject to its ruling, which may include
  disqualification.
 \item
  Campaigning shall be defined as, but not limited to:

  \begin{enumerate}
   \item
    Distribution of campaign materials;
   \item
    Speaking to classes, residences, student groups, or individuals for
    the purpose of presenting a platform as a candidate for a position;
   \item
    Sharing of website links or social media pages that are relevant to
    campaigns;
   \item
    Any action performed primarily for the purpose of seeking election.
  \end{enumerate}
 \item
  Campaign material is defined as, but is not limited to:

  \begin{enumerate}
   \item
    Campaign posters
   \item
    Campaign swag/apparel
   \item
    Social media pages or events, including any relevant posts, photos
    or documents contained therein
   \item
    Personal campaign websites
  \end{enumerate}
 \item
  To promote the election and increase voter turnout, all campaign
  material must include:

  \begin{enumerate}
   \item
    the MES logo
   \item
    the date and location of polling
   \item
    the position for which the candidate is seeking election
  \end{enumerate}
 \item
  All campaigning shall be in good taste (no promoting violence,
  substance abuse or discrimination) and fairly conducted with courtesy
  to other candidates. No slander or libel will be tolerated.

\end{enumerate}

\hypertarget{closure-of-campaign-period}{%
 \subsubsection{Closure of Campaign Period}
 \label{closure-of-campaign-period}}
\begin{enumerate}
 \item
  Candidates are to cease campaigning once the campaign period has ended
  and voting has begun.
 \item
  Once the campaign period has ended, no more campaign material may be
  posted by the candidate or their campaign team. Any campaign materials
  posted prior to the closing of the campaign period may remain visible
  and available in order to keep voters informed of candidates and their
  intentions. The candidates are still expected to abide by the campaign
  rules, determined by the Elections Committee.
 \item
  Campaign posters must be removed within 24 hours after the polls
  close.
 \item
  9.3) Campaign Team
 \item
  Candidates choosing to recruit a campaign team must submit names and
  emails of all members of their campaign team to the Elections
  Committee.
 \item
  Campaign team members are the responsibility of the candidate, and
  must follow all election rules and procedures.
 \item
  At the discretion of the Elections Committee, MES members that are
  actively campaigning for a specific candidate will be added to that
  candidate's team.

\end{enumerate}

\hypertarget{campaign-posters}{%
 \subsubsection{Campaign Posters}
 \label{campaign-posters}}
\begin{enumerate}
 \item
  The content and placement of the campaign posters should follow MSU
  guidelines.
 \item
  The Chief Returning Officer must vet and approve all campaign posters
  to ensure the required information is included.
 \item
  The Chief Returning Officer shall be responsible for obtaining the MSU
  Underground's approval of all campaign posters.
 \item
  Only masking tape, regular staples and tacks are used to put up
  posters on appropriate boards. Only masking tape shall be used to put
  up posters on appropriate walls. Duct tape, packing tape, glue, staple
  guns, etc. are prohibited.
 \item
  Candidates may not use physical posters or banners that exceed 11'' x
  17'' inches.
 \item
  Upon submission of receipts, up to a maximum of \$5 will be reimbursed
  to candidates. This is to ensure that all candidates have equal
  opportunity to campaign via printed posters.
\end{enumerate}

\hypertarget{online-and-social-media-campaigning}{%
 \subsubsection{Online and Social Media Campaigning}
 \label{online-and-social-media-campaigning}}
\begin{enumerate}
 \item
  All digital campaign materials must follow the guidelines of campaign
  conduct.
 \item
  The Elections Committee may modify the rules of social media
  campaigning, as long as notice is duly given to candidates and the
  bylaws are amended accordingly at subsequent Council meetings.
 \item
  Digital campaigning must be carried out in a manner where candidates
  do not spam or otherwise make their candidacy announced without direct
  interaction from potential voters. Digital campaigning must also
  minimize the leverage of a candidate's existing connections and
  networks. This includes, but is not limited to, the following:

  \begin{enumerate}
   \item
    Creating new, alternate, or throwaway accounts
   \item
    Modifying the handle, tag, username, status, or equivalent to refer
    to their candidacy
   \item
    Receiving endorsements or having posts reshared by MES Groups
  \end{enumerate}
 \item
  Digital campaigning must be carried out in a manner where candidates
  do not campaign in group chats or any situation where it would be
  considered spam if all candidates proceeded to campaign in a similar
  manner. This would include, but is not limited to:

  \begin{enumerate}
   \item
    Announcing candidacy on Reddit, Discord, or other platforms
   \item
    Announcing candidacy in academic or social group chats
  \end{enumerate}
 \item
  Approved digital campaigning includes, but is not limited to the
  following:

  \begin{enumerate}
   \item
    Creating a personal webpage or website
   \item
    Changing the viewability of your account to public
   \item
    Modifying the bio, about me, or equivalent section of profile, given
    that potential voters cannot view this without directly navigating
    to your profile.
   \item
    Sharing posts created by the MES
   \item
    Creating posts, memes, stories, reels, videos, or other content
   \item
    Asking Discord server moderators to announce MES Elections are
    happening, given there is no acknowledgement of the candidate
  \end{enumerate}
 \item
  The Elections Committee can approve any digital campaign materials by
  request of a candidate. Candidates are encouraged to err on the side
  of caution.

\end{enumerate}

\hypertarget{other-campaign-procedures}{%
 \subsubsection{Other Campaign Procedures}
 \label{other-campaign-procedures}}
\begin{enumerate}
 \item
  Presidential candidates may submit a statement no longer than 500
  words to the Chief Returning Officer by the end of the first day of
  the campaign period. These statements will be posted on the MES
  website and on other MES social media.
 \item
  Candidates in general elections may submit a statement no longer than
  250 words to the Chief Returning Officer by the end of the first day
  of the campaign period. These statements will be posted on the MES
  website and on other MES social media.
 \item
  Candidates must submit all expenses made during the election to the
  Chief Returning Officer, with receipts. The total amount of campaign
  expenses per candidate may not exceed \$50 during any election.
  Candidates shall not be reimbursed for any campaign expenses, except
  for the \$5 poster printing stipend, as outlined above in section
  9.4).
 \item
  When candidates are campaigning in lectures they shall consult the
  professor prior to the class and must get permission to do the
  campaign presentation.

\end{enumerate}

\hypertarget{elections-committee}{%
 \subsection{Elections Committee}
 \label{elections-committee}}
\begin{enumerate}
 \item
  The Elections Committee shall consist of the Chief Returning Officer
  as a non-voting chair, , the Equity and Inclusion Officer, the VPSL,
  and the President. If any member of the Elections Committee is seeking
  office in an MES election or demonstrates an obvious bias to any one
  candidate, the MES Executive shall appoint an unbiased MES Council
  member who is not seeking election.
 \item
  The Elections Committee shall meet prior to the first of September of
  each year to set the dates for all elections. These dates shall be
  made ready for advertisement to all McMaster undergraduate students in
  the Faculty of Engineering once the dates are set.
 \item
  For all MES elections, the Elections Committee shall:

  \begin{enumerate}
   \item
    Post a list of all available positions to be filled and a summary of
    their duties prior to the nomination period.
   \item
    Post all election notices on the MES website, and on other social
    and physical media.
   \item
    Review and update the campaign rules and inform candidates of the
    campaign rules during their respective nomination periods.
   \item
    Pass judgment as to the eligibility of all the candidates (see MES
    Bylaws Section C.1.1), and ensure that correct nomination procedures
    have been followed.
   \item
    Ensure that all campaign materials are in accordance with MSU
    regulations.
   \item
    Not engage in the campaign of any candidate(s).
   \item
    Be responsible for drawing up the official ballot for all MES
    elections.
   \item
    Be responsible for running all elections in accordance with the MES
    Bylaws Section C.
   \item
    Post the names of all successful candidates as soon as they are
    known. Except for presidential candidates, which will be announced
    at Fireball.
   \item
    Take any corrective or disciplinary actions necessary with regard to
    electoral rule violations of any type, as outlined in this or any
    other duly enacted document, with the provision that the MES
    Executive, excepting any members thereof directly affected by the
    matter in question, shall act as the final arbitrator in all
    disputes.
   \item
    Prepare a post-election report detailing the precise results of the
    election and the nature of actions taken in regard to any protests,
    disputes, and/or rule contraventions during the course of the
    election. Such a report must be prepared after every MES election
    and submitted to the MES Council.
  \end{enumerate}
\end{enumerate}

\hypertarget{election-procedures}{%
 \subsection{Election Procedures}
 \label{election-procedures}}
\begin{enumerate}
 \item
  Voting for elections may be conducted online. If elections are held
  online and more than one candidate is running for the same position, a
  single transferable voting system shall be used.
 \item
  An election shall be declared invalid if:

  \begin{enumerate}
   \item
    The Elections Committee rules that the number and seriousness of
    election rule and procedure infractions that may have affected the
    election results warrants the invalidation of the election.
  \end{enumerate}
 \item
  There must be both a non-confidence and an abstain voting option on
  the ballots of all MES elections.
 \item
  The candidate receiving the greatest number of valid votes for a given
  position shall be declared elected to that position, subject to
  ratification.
 \item
  In the case of a tie, the Chief Returning Officer shall reopen the
  campaign period for the tied position for a period of at least four
  days. Only those candidates who tied will be allowed to campaign
  again. An additional voting day, immediately following the second
  campaign period, will be opened up to all McMaster undergraduate
  students in the Faculty of Engineering of the constituency in question
 \item
  In the event that a member of the MES will not be able to vote on the
  day of an election, they may vote in advance. It is the responsibility
  of the voter to make proper arrangements as listed below:

  \begin{enumerate}
   \item
    An advance vote must be made in the presence of the Chief Returning
    Officer and one other Elections Committee member.
   \item
    After the vote is completed, the Chief Returning Officer and
    Elections Committee member will both initial the ballot and place it
    in the advance poll envelope.
   \item
    Any ballot placed in the advance poll envelope shall not be
    interfered with in any way. The ballots must only be removed from
    the envelope while the entire Elections Committee is present, to add
    the votes as recorded on the ballots to the election results.
   \item
    The advance poll envelope shall be the responsibility of the Chief
    Returning Officer
  \end{enumerate}
\end{enumerate}