\subsubsection{Vice President, Student
  Life}
\label{vice-president-student-life}
The Vice President, Student Life shall:

\begin{enumerate}
 \item
  Act as the President in all capacities in the latter's absence.
 \item
  Serve as a member of the following committees:

  \begin{enumerate}
   \item
    Elections Committee (see MES Bylaws Section \ref{elections-committee})
   \item
    Awards Selection Committee (see MES Bylaws Section \ref{awards-selection-committee})
   \item
    Sports Committee (see MES Bylaws Section \ref{sports-committee})
   \item
    First Year Committee (see MES Bylaws Section \ref{first-year-committee})
   \item
    Hatch Operations Committee (Faculty committee)
  \end{enumerate}
 \item
  Be an ex-officio member of all MES Committees except those to which
  they have been duly appointed.
 \item
  Assist the VP Academic in the planning of Back2School Week (the week
  following Welcome Week), and shall organize Hatch Open House as part
  of Back2School Week with assistance from the AVPC.
 \item
  Coordinate and organize events for Frost Week.
 \item
  Facilitate collection of content for the MESsenger, and pass it on to
  VPC to send to all undergraduate engineering students on a monthly
  basis
 \item
  Supervise the activities and initiatives of the Program and First Year
  Representatives.
 \item
  Plan an end-of-semester social for all MES appointed and elected
  council members.
 \item
  Be responsible for approving any interfaculty morale boosting
  activities on behalf of the MES.
 \item
  Liaise with Environmental and Occupational Health and Safety Services
  and ensure all MES social activities abide by its guidelines.
 \item
  Provide the Administrator with dates and details of all events,
  conferences, and competitions.
 \item
  Chair the First Year Committee and act as an advisor to the First Year
  Representatives to help coordinate activities and fundraising
  initiatives (see MES Bylaws Section \ref{first-year-committee}).
 \item
  Report to the President.
 \item
  Supervise the following positions:

  \begin{enumerate}
   \item
    Associate Vice-President Events
   \item
    Wellness Coordinator(s)
   \item
    Bus Monitor Lead
   \item
    First Year Representatives
   \item
    Sports Coordinator(s)
   \item
    Sustainability Coordinator(s)
  \end{enumerate}
 \item
  Develop protocols, with any members at the VPSL's discretion, for the
  purposes of risk aversion and incident mitigation for all aspects of
  MES events. To be completed by September.

\end{enumerate}