\section{Awards}
\label{awards}

\subsection{General Policies}
\label{general-policies}
\begin{enumerate}
 \item
  All annual awards and their descriptions must be listed in the
  Engineering Handbook.
 \item
  General Nomination Procedure:

  \begin{enumerate}
   \item
    Notice of all nomination periods and procedures must be clearly
    posted on the MES website and any other social media, and published
    in the Frequency.
   \item
    All nomination forms must include the name and contact information
    of the nominee and the award for which they are being nominated.
   \item
    Space must be available for comments outlining the reasons for
    nomination.
   \item
    All nomination forms received after the close of the nomination
    period shall be declared invalid with no exception.
  \end{enumerate}
 \item
  The Awards Selection Committee will select the recipients of MES
  awards in accordance with the criteria of each award (see MES Bylaws
  Section \ref{awards-selection-committee}).
 \item
  In order to create new awards, whether they be recurring or unique:

  \begin{enumerate}
   \item
    All proposals must have established the terms of the fund and the
    criterion by which candidates will be judged. A statement of the
    terms of the award must be added to the Policy Manual. % clerical - soon to be nonexistent
   \item
    The method of administering the awards must be decided upon prior to
    passage of the award.
   \item
    New awards to be funded by the MES must be passed by a majority vote
    at an MES Executive meeting for amounts up to \$500, an MES Council
    meeting for amounts up to \$1,000, or an MES General Meeting for
    amounts exceeding \$1,000.
   \item
    The MES Council must approve awards funded by an external source,
    through a two thirds majority vote at an MES Council meeting.
  \end{enumerate}
\end{enumerate}

\subsection{MES Awards}
\label{mes-awards}
\begin{enumerate}
 \item
  Council Appreciation Award

  \begin{enumerate}
   \item
    wUp to three Council Appreciation Awards shall be given out each
    year at Fireball for exceptional work on the MES Council.
   \item
    Recipients must be a member of the MES Council, and may be either
    appointed or elected.
   \item
    MES Executive members are ineligible for this award.
   \item
    The MES Executive shall choose the recipients for this award..
  \end{enumerate}
 \item
  ``Image of an Engineer'' Award

  \begin{enumerate}
   \item
    Up to three ``Image of an Engineer'' Awards shall be given out each
    year at Fireball.
   \item
    The ``Image of an Engineer'' Award is given to an undergraduate
    Faculty of Engineering student who contributes to improving the
    image of an engineer through extracurricular, leadership, and/or
    community activities.
  \end{enumerate}
 \item
  Faculty Appreciation Award

  \begin{enumerate}
   \item
    Up to two Faculty Appreciation Awards shall be given out each year
    at Fireball.
   \item
    The Faculty Appreciation Award is given to a faculty or staff member
    within the Faculty of Engineering, to recognize a significant
    contribution outside of the classroom to the undergraduate Faculty
    of Engineering students at McMaster University.
  \end{enumerate}
 \item
  Outstanding Teaching Assistant Award

  \begin{enumerate}
   \item
    Up to two Outstanding Teaching Assistant Awards shall be given out
    each year at Fireball.
   \item
    The Outstanding Teaching Assistant Award is given to a Teaching
    Assistant, of a Faculty of Engineering course, who has gone above
    and beyond in order to provide an engaging learning environment for
    their students.
  \end{enumerate}
 \item
  President's Award

  \begin{enumerate}
   \item
    Up to two President's Awards shall be given out each year at
    Fireball.
   \item
    The President's Award is given to an MES member for their
    exceptional contributions to the MES.
   \item
    The President is ineligible for this award.
   \item
    The President shall choose the recipients for this award.
  \end{enumerate}
\end{enumerate}
